Effectiveness of visual test in linear regression diagnositics

### 2022-07-28 paper_comparison/paper_comparison/paper_comparison.Rmd
draft model fitting

As suggested by many previous studies [many studies refs here], it is reasonable for an subject to evaluate a block of 20 lineups. Subjects for both two experiments were recruited from Prolific web service [ref here]. Data screening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate and 10 minimum submissions. During the experiment, they were asked to select one or more plots that are mostly different from others, provide a reason for their selections, and express how different they think the selected plots are from others. If there was no noticeable difference between plots in a lineup, subject could choose to select zero plots without providing the reason. Information about gender, age, education, and previous experience in visual experiment were also collected. No subject was shown the same lineup twice.

@@

VI_MODEL$plot(ori_dat) + ggtitle(quote(He[2]:"U shape")) -> p1
VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote(He[3]:"S shape")) -> p2
VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote(He[6]:"M shape")) -> p3
VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote(He[18]:"Triple-U shape")) -> p4

@@

VI_MODEL$plot(ori_dat) + ggtitle(quote(sigma~"="~0.5)) -> p1
VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
  ggtitle(quote(sigma~"="~1)) -> p2
VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
  ggtitle(quote(sigma~"="~2)) -> p3
VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
  ggtitle(quote(sigma~"="~4)) -> p4

@@

VI_MODEL$plot(ori_dat) + ggtitle(quote(U(-1,1))) -> p1
VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle(quote(N(0,0.3^2))) -> p2
VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle(quote(lognormal(0,0.6^2)/3)) -> p3
VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle(quote(U~"{1,5}")) -> p4

@@

@@

[table ref here] tabulates the number of subjects, genders, and lineups evaluated after applying the data screening procedure.

```{r}
polynomials %>%
  group_by(set) %>%
  summarise(across(age_group:previous_experience, first)) %>%
  mutate(education = fct_relevel(education, "High School or below")) %>%
  ggplot() +
  geom_bar(aes(y = education), orientation = "y") -> education_plot

@@

For each parameter combination, effect $E$ is derived as $E = \frac{1}{2}(\boldsymbol{R}_a\boldsymbol{X}_b)'diag(\sigma^2\boldsymbol{R}_a)\boldsymbol{R}_a\boldsymbol{X}_b$ from the Kullback-Leibler divergence. [KL divergence ref here] (The derivation is provided in [appendix ref here].). The logistic regression is fit using $E$ as the only fixed effect covariate

@@

# Data processing

>>>>

As suggested by many previous studies [many studies refs here], it is reasonable for an subject to evaluate a block of 20 lineups. Subjects for both two experiments were recruited from Prolific web service [ref here]. Data screening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate and 10 minimum submissions. During the experiment, they were asked to select one or more plots that are mostly different from others, provide a reason for their selections, and express how different they think the selected plots are from others. If there was no noticeable difference between plots in a lineup, subject could choose to select zero plots without providing the reason. Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected. No subject was shown the same lineup twice.

@@

VI_MODEL$plot(ori_dat) + 
  ggtitle(quote(He[2]:"U shape")) + 
  theme_light() -> p1

VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote(He[3]:"S shape")) + 
  theme_light() -> p2

VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote(He[6]:"M shape")) + 
  theme_light() -> p3

VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote(He[18]:"Triple-U shape")) + 
  theme_light() -> p4

@@

VI_MODEL$plot(ori_dat) + 
  ggtitle(quote(sigma~"="~0.5)) + 
  theme_light() -> p1

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
  ggtitle(quote(sigma~"="~1)) + 
  theme_light() -> p2

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
  ggtitle(quote(sigma~"="~2)) + 
  theme_light() -> p3

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
  ggtitle(quote(sigma~"="~4)) + 
  theme_light() -> p4

@@

VI_MODEL$plot(ori_dat) + 
  ggtitle(quote(U(-1,1))) + 
  theme_light() -> p1

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle(quote(N(0,0.3^2))) + 
  theme_light() -> p2

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle(quote(lognormal(0,0.6^2)/3)) + 
  theme_light() -> p3

VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle(quote(U~"{1,5}")) + 
  theme_light() -> p4

@@

In overall, there were a total of 3200 lineup evaluations made by 160 subjects in experiment I, where 320 lineup evaluations were attention checks and were not used in the following analysis. The collated dataset is provided in `polynomials` of the `visage` `R` package.

@@

Table \ref{tab:demographic-table} tabulates the number of subjects, preferred pronouns, education backgrounds, age groups, and previous experience in visual experiment. Figure \ref{fig:demographic-summary-plot} visulizes the marginal distribution of each of the category. The collated data was a balanced sample among male and female. Most of the ages of subject were between 18 to 39, while many of them were between 18 to 24. Very few subjects were awarded degree higher than Bachelor Degree. Around 40% of subjects had previous experience in visual experiment.

```{r results="asis"}

polynomials %>%
  group_by(set) %>%
  slice_sample(n = 1) %>%
  ungroup() %>%
  count(pronoun, education, age_group, previous_experience) %>%
  arrange(desc(n)) %>%
  knitr::kable(col.names = c("Preferred pronoun", "Education", "Age group", "Previous experience", "Subject"),
               escape = FALSE,
               format = "latex",
               linesep = "",
               booktabs = TRUE, 
               caption  = 'Summary of demographic information', 
               label = "demographic-table" ) %>%
  cat()
```

```{r demographic-summary-plot, fig.cap="Summary of demographic information. \\label{fig:demographic-summary-plot}"}
polynomials %>%
  group_by(set) %>%
  summarise(across(age_group:previous_experience, first)) %>%
  mutate(education = fct_relevel(education, "High School or below")) %>%
  ggplot() +
  geom_bar(aes(y = education), orientation = "y") -> education_plot

@@

For each parameter combination, effect $E$ is derived from the Kullback-Leibler divergence (the derivation is provided in [appendix ref here]) formulated as

\begin{equation} \label{eq:effect-size-ex1}
E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
\end{equation}

where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.

The logistic regression is fit using $log_e(E)$ as the only fixed effect covariate for the power of visual test formulated as

\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
\end{equation}

where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.

Table [table ref here] shows the parameter estimates of the logistic regression. [Discussion about the numeric estimates here]

## Power comparison

Figure [figure ref here] shows an overview of estimated power against natural logarithm

@@

# Old

### 2022-07-28 paper_comparison/paper_comparison/paper_comparison.tex
draft model fitting

As suggested by many previous studies {[}many studies refs here{]}, it
is reasonable for an subject to evaluate a block of 20 lineups. Subjects
for both two experiments were recruited from Prolific web service {[}ref
here{]}. Data screening procedure was applied during the recruitment,
subjects were required to be fluent in English, with \(98\%\) minimum
approval rate and 10 minimum submissions. During the experiment, they
were asked to select one or more plots that are mostly different from
others, provide a reason for their selections, and express how different
they think the selected plots are from others. If there was no
noticeable difference between plots in a lineup, subject could choose to
select zero plots without providing the reason. Information about
gender, age, education, and previous experience in visual experiment
were also collected. No subject was shown the same lineup twice.

@@

@@

{[}table ref here{]} tabulates the number of subjects, genders, and
lineups evaluated after applying the data screening procedure.

\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-3-1.pdf}

@@

For each parameter combination, effect \(E\) is derived as
\(E = \frac{1}{2}(\boldsymbol{R}_a\boldsymbol{X}_b)'diag(\sigma^2\boldsymbol{R}_a)\boldsymbol{R}_a\boldsymbol{X}_b\)
from the Kullback-Leibler divergence. {[}KL divergence ref here{]} (The
derivation is provided in {[}appendix ref here{]}.). The logistic
regression is fit using \(E\) as the only fixed effect covariate

\hypertarget{data-processing}{%
\section{Data processing}\label{data-processing}}

\hypertarget{results-1}{%
\section{Results}\label{results-1}}

>>>>

As suggested by many previous studies {[}many studies refs here{]}, it
is reasonable for an subject to evaluate a block of 20 lineups. Subjects
for both two experiments were recruited from Prolific web service {[}ref
here{]}. Data screening procedure was applied during the recruitment,
subjects were required to be fluent in English, with \(98\%\) minimum
approval rate and 10 minimum submissions. During the experiment, they
were asked to select one or more plots that are mostly different from
others, provide a reason for their selections, and express how different
they think the selected plots are from others. If there was no
noticeable difference between plots in a lineup, subject could choose to
select zero plots without providing the reason. Information about
preferred pronoun, age group, education, and previous experience in
visual experiment were also collected. No subject was shown the same
lineup twice.

@@

In overall, there were a total of 3200 lineup evaluations made by 160
subjects in experiment I, where 320 lineup evaluations were attention
checks and were not used in the following analysis. The collated dataset
is provided in \texttt{polynomials} of the \texttt{visage} \texttt{R}
package.

@@

Table \ref{tab:demographic-table} tabulates the number of subjects,
preferred pronouns, education backgrounds, age groups, and previous
experience in visual experiment. Figure
\ref{fig:demographic-summary-plot} visulizes the marginal distribution
of each of the category. The collated data was a balanced sample among
male and female. Most of the ages of subject were between 18 to 39,
while many of them were between 18 to 24. Very few subjects were awarded
degree higher than Bachelor Degree. Around 40\% of subjects had previous
experience in visual experiment.

\caption{\label{tab:demographic-table}Summary of demographic information}
\centering
\begin{tabular}[t]{llllr}
\toprule
Preferred pronoun & Education & Age group & Previous experience & Subject\\
\midrule
She & Diploma and Bachelor Degree & 18-24 & No & 17\\
He & Diploma and Bachelor Degree & 18-24 & Yes & 14\\
She & Diploma and Bachelor Degree & 18-24 & Yes & 12\\
He & Diploma and Bachelor Degree & 18-24 & No & 11\\
He & Diploma and Bachelor Degree & 25-39 & No & 10\\
She & Diploma and Bachelor Degree & 25-39 & No & 10\\
She & Diploma and Bachelor Degree & 25-39 & Yes & 9\\
He & High School or below & 18-24 & No & 8\\
She & High School or below & 18-24 & No & 7\\
He & High School or below & 18-24 & Yes & 6\\
He & Diploma and Bachelor Degree & 25-39 & Yes & 5\\
He & Masters Degree & 25-39 & No & 5\\
He & Masters Degree & 25-39 & Yes & 5\\
He & High School or below & 25-39 & No & 4\\
She & High School or below & 25-39 & No & 4\\
She & Masters Degree & 25-39 & Yes & 4\\
He & High School or below & 25-39 & Yes & 3\\
She & High School or below & 18-24 & Yes & 3\\
She & Masters Degree & 25-39 & No & 3\\
He & Honours Degree & 25-39 & No & 2\\
She & Diploma and Bachelor Degree & 40-54 & No & 2\\
She & Honours Degree & 18-24 & No & 2\\
He & High School or below & 40-54 & Yes & 1\\
He & High School or below & 55-64 & No & 1\\
He & Masters Degree & 40-54 & Yes & 1\\
He & Masters Degree & 55-64 & No & 1\\
Other & High School or below & 25-39 & No & 1\\
She & High School or below & 40-54 & No & 1\\
She & Honours Degree & 25-39 & No & 1\\
She & Honours Degree & 25-39 & Yes & 1\\
She & Masters Degree & 18-24 & No & 1\\
She & Masters Degree & 40-54 & No & 1\\
They & Diploma and Bachelor Degree & 18-24 & No & 1\\
They & Diploma and Bachelor Degree & 25-39 & No & 1\\
They & High School or below & 18-24 & No & 1\\
They & High School or below & 25-39 & No & 1\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{paper_comparison_files/figure-latex/demographic-summary-plot-1.pdf}
\caption{Summary of demographic information.
\label{fig:demographic-summary-plot}}
\end{figure}

@@

For each parameter combination, effect \(E\) is derived from the
Kullback-Leibler divergence (the derivation is provided in {[}appendix
ref here{]}) formulated as

\begin{equation} \label{eq:effect-size-ex1}
E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
\end{equation}

where \(diag(.)\) is the diagonal matrix constructed from the diagonal
elements of \(\boldsymbol{R}_a\).

The logistic regression is fit using \(log_e(E)\) as the only fixed
effect covariate for the power of visual test formulated as

\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
\end{equation}

where \(\Lambda(.)\) is the standard logistic function given as
\(\Lambda(z) = exp(z)/(1+exp(z))\).

Table {[}table ref here{]} shows the parameter estimates of the logistic
regression. {[}Discussion about the numeric estimates here{]}

\hypertarget{power-comparison}{%
\subsection{Power comparison}\label{power-comparison}}

Figure {[}figure ref here{]} shows an overview of estimated power
against natural logarithm

\hypertarget{old}{%
\section{Old}\label{old}}
